Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2019-06-11 thirdPaper/draft/typeConeAssociahedra.tex
Improved Prop.4.45 and Coro 4.46.

The extriangulated category~$K^{[-1,0]}(\proj \Lambda)$ is the setting in which we can finally prove point~(ii) of Proposition~\ref{prop:exchangeablePairsNKC}, using results of~\cite{DemonetIyamaJasso} on~$\b g$-vectors.  
% Proposition~\ref{prop::2kbproj-is-extriangulated} gives a setting in which to apply Theorem~\ref{}.
Before doind so, we introduce the notion of mutation conflation for two-term silting objects.

@@

\begin{definition}\label{defn::mutation conflation}
A conflation $X\infl E\defl Y$ of~$K^{[-1,0]}(\proj \Lambda)$ is called a \defn{mutation conflation} if there are basic, two-term, silting objects $X\oplus R$, $Y\oplus R$ such that the inflation $X\infl E$ is a left ($\add R$)-approximation (i.e. any morphism from~$X$ to an object in~$\add R$ factors through~$X\infl E$).
\end{definition}

\begin{proposition}
\label{prop::exchange conflations are unique}
 Let~$\Lambda$ be a finite-dimensional~$\field$-algebra, where~$\field$ is a field.  Let~$X$ and~$Y$ be objects of~$K^{[-1,0]}(\proj \Lambda)$.  Assume that~$X\infl E\defl Y$ and~$X\infl E'\defl Y$ (or~$Y\infl E'\defl X$) are two mutation conflations (see Definition~\ref{defn::mutation conflation}.  
 Then~$E$ and~$E'$ are isomorphic. \pg{Is it possible that~$X$ and~$Y$ are permuted in one of the exchange conflations?}
\end{proposition}
\begin{proof}
 By~\cite[Theorem 6.5]{DemonetIyamaJasso} (based on~\cite[Theorem 2.3]{DehyKeller}), rigid objects in~$K^{[-1,0]}(\proj \Lambda)$ are determined by their~$\b g$-vector.  The datum of the~$\b g$-vector of an object of~$K^{[-1,0]}(\proj \Lambda)$ is equivalent to that of its class in the Grothendieck group~$K_0(K^{[-1,0]}(\proj \Lambda))$, where~$K^{[-1,0]}(\proj \Lambda)$ is viewed as an extriangulated category (see Proposition~\ref{prop::2kbproj-is-extriangulated}).  In this Grothendieck group, we have that~$[E] = [E'] = [X]+[Y]$.  Thus~$E$ and~$E'$ are isomorphic.
\end{proof}

\begin{corollary}\label{coro::exchangeablePairsNKC}
 Point~(ii) of Proposition~\ref{prop:exchangeablePairsNKC} holds.
\end{corollary}
\begin{proof}
 Let~$\omega$ and~$\omega'$ be in facets~$F$ and~$F'$ as in the statement of Proposition~\ref{prop:exchangeablePairsNKC}.  For every walk~$\omega''$, let~$P(\omega'')$ be the object of~$K^{[-1,0]}(\proj \Lambda)$ corresponding to~$\omega''$.  Then the walks~$\mu$ and~$\nu$ defined in the same statement yield an exchange conflation
 \[
  P(\omega) \infl P(\mu) \oplus P(\nu) \defl P(\omega') \quad \textrm{or} \quad P(\omega') \infl P(\mu) \oplus P(\nu) \defl P(\omega).
 \]
 Assume that there are other facets~$F_0$ and~$F_0'$ also satisfying the hypotheses of Proposition~\ref{prop:exchangeablePairsNKC}.  Then they give rise to walks~$\mu'$ and~$\nu'$ and an exchange conflation
 \[
  P(\omega) \infl P(\mu') \oplus P(\nu') \defl P(\omega') \quad \textrm{or} \quad P(\omega') \infl P(\mu') \oplus P(\nu') \defl P(\omega).
 \]
 By Proposition~\ref{prop::exchange conflations are unique}, we get that~$P(\mu)\oplus P(\nu)$ and~$P(\mu')\oplus P(\nu')$ are isomorphic.  This implies that~$\{\mu,\nu\}$ and~$\{\mu',\nu'\}$ are the same sets.  This determines~$\sigma$, and point (ii) of Proposition~\ref{prop:exchangeablePairsNKC} is proved.
\end{proof}

@@

\subsubsection{A quotient of a Frobenius category}
\begin{example}
 Consider the preprojective algebra of type~$A_3$, as in Example~\ref{exam::preprojective silting objects}.  It is a self-injective algebra, so its module category is a Frobenius category.  Its projective-injective objects are the modules 
 \[
  \begin{matrix} 1 \\ 2 \\ 3 \end{matrix} \quad , \quad  \begin{matrix} 2 \\ 13 \\ 2 \end{matrix} \quad \textrm{and} \quad \begin{matrix} 3 \\ 2 \\ 1 \end{matrix}.
 \]
 Taking the quotient by the ideal generated by all morphisms factoring through direct sums of copies of the second of these objects, we get an extriangulated category whose Auslander--Reiten quiver is given by
  \[
  \hspace*{-2cm}
 {\footnotesize
  \begin{tikzcd}
   \begin{matrix} 3 \end{matrix}\arrow[dr] && \begin{matrix} 2 \\ 1 \end{matrix}\arrow[r]\arrow[dr] & \begin{matrix} 3 \\ 2 \\ 1 \end{matrix}\arrow[r]& \begin{matrix} 3 \\ 2 \end{matrix}\arrow[dr] && \begin{matrix} 1 \end{matrix}\arrow[dr]&& \begin{matrix} 2 \\ 3 \end{matrix} \\
    & \begin{matrix} 2\\ 13 \end{matrix}\arrow[ur]\arrow[dr] && \begin{matrix} 2 \end{matrix}\arrow[ur]\arrow[dr] && \begin{matrix} 13\\2 \end{matrix}\arrow[ur]\arrow[dr] && \begin{matrix} 2\\13 \end{matrix}\arrow[ur]\arrow[dr] &  \\
   \begin{matrix} 1 \end{matrix}\arrow[ur] && \begin{matrix} 2\\3 \end{matrix}\arrow[ur]\arrow[r] & \begin{matrix} 1\\2\\3 \end{matrix}\arrow[r] & \begin{matrix} 1\\2 \end{matrix}\arrow[ur] && \begin{matrix} 3\\1 \end{matrix}\arrow[ur] && \begin{matrix} 2\\1 \end{matrix}
  \end{tikzcd}
  }
 \]
\end{example}

>>>>

The extriangulated category~$K^{[-1,0]}(\proj \Lambda)$ is the setting in which we can finally prove point~(ii) of Proposition~\ref{prop:exchangeablePairsNKC}, using results of~\cite{DemonetIyamaJasso} on~$\b g$-vectors.  
% Proposition~\ref{prop::2kbproj-is-extriangulated} gives a setting in which to apply Theorem~\ref{}.
Before doing so, we introduce the notion of mutation conflation for two-term silting objects.

@@

\begin{definition}\label{defn::mutation conflation}
A conflation $X\infl E\defl Y$ of~$K^{[-1,0]}(\proj \Lambda)$ is called a \defn{mutation conflation} if there are basic, two-term, silting objects $X\oplus R$, $Y\oplus R$ such that the inflation $X\infl E$ is a left ($\add R$)-approximation (i.e. any morphism from~$X$ to an object in~$\add R$ factors through~$X\infl E$). \pg{To simplify, we should perhaps ask for the approximations to be minimal.}
\end{definition}

\begin{proposition}
\label{prop::exchange conflations are unique}
 Let~$\Lambda$ be a finite-dimensional~$\field$-algebra, where~$\field$ is a field.  Let~$X$ and~$Y$ be objects of~$K^{[-1,0]}(\proj \Lambda)$.  
 \begin{enumerate}
  \item If there is a mutation conflation (see Definition~\ref{defn::mutation conflation}) of the form~$X\infl E\defl Y$, then there can be no mutation conflation of the form~$Y\infl E'\defl X$.

  \item Assume that~$X\infl E\defl Y$ and~$X\infl E'\defl Y$ are two mutation conflations.  
 Then~$E$ and~$E'$ are isomorphic. 
 \end{enumerate} 
\end{proposition}
\begin{proof}
 We first prove~(1).  Assume that~$X\infl E\defl Y$ and~$Y\infl E'\defl X$ are two mutation conflations.  Applying the functor~$\Hom{K^b}(\susp^{-1} Y, -)$ to the second sequence, we get an exact sequence of abelian groups
 \[
  \Hom{K^b}(\susp^{-1} Y, E') \to \Hom{K^b}(\susp^{-1} Y, X) \Hom{K^b}(\susp^{-1} Y, \susp Y).
 \]
 By definition of a mutation conflation, we have that~$\Hom{K^b}(\susp^{-1} Y, E') = 0$.  Moreover, since~$Y$ is concentrated in homological degrees~$-1$ and~$0$, we get that~$\Hom{K^b}(\susp^{-1} Y, \susp Y) = 0$.  Therefore,~$\Hom{K^b}(\susp^{-1} Y, X) = 0$. 
 However, since the conflation~$X\infl E\defl Y$ is not split, we have that~$\Hom{K^b}(\susp^{-1} Y, X) \neq 0$, a contradiction.  This proves~(1).

 \smallskip

 We now prove~(2).  Let~$X\infl E\defl Y$ and~$X\infl E'\defl Y$ be two mutation conflations.  By~\cite[Theorem 6.5]{DemonetIyamaJasso} (based on~\cite[Theorem 2.3]{DehyKeller}), rigid objects in~$K^{[-1,0]}(\proj \Lambda)$ are determined by their~$\b g$-vector.  The datum of the~$\b g$-vector of an object of~$K^{[-1,0]}(\proj \Lambda)$ is equivalent to that of its class in the Grothendieck group~$K_0(K^{[-1,0]}(\proj \Lambda))$, where~$K^{[-1,0]}(\proj \Lambda)$ is viewed as an extriangulated category (see Proposition~\ref{prop::2kbproj-is-extriangulated}).  In this Grothendieck group, we have that~$[E] = [E'] = [X]+[Y]$.  Thus~$E$ and~$E'$ are isomorphic.
\end{proof}

\begin{corollary}\label{coro::exchangeablePairsNKC}
 Point~(ii) of Proposition~\ref{prop:exchangeablePairsNKC} holds.
\end{corollary}
\begin{proof}
 Let~$\omega$ and~$\omega'$ be in facets~$F$ and~$F'$ as in the statement of Proposition~\ref{prop:exchangeablePairsNKC}.  For every walk~$\omega''$, let~$P(\omega'')$ be the object of~$K^{[-1,0]}(\proj \Lambda)$ corresponding to~$\omega''$.  Then the walks~$\mu$ and~$\nu$ defined in the same statement yield a mutation conflation
 \[
  P(\omega) \infl P(\mu) \oplus P(\nu) \defl P(\omega') \quad \textrm{or} \quad P(\omega') \infl P(\mu) \oplus P(\nu) \defl P(\omega).
 \]
 Assume that there are other facets~$F_0$ and~$F_0'$ also satisfying the hypotheses of Proposition~\ref{prop:exchangeablePairsNKC}.  Then they give rise to walks~$\mu'$ and~$\nu'$ and an exchange conflation
 \[
  P(\omega) \infl P(\mu') \oplus P(\nu') \defl P(\omega') \quad \textrm{or} \quad P(\omega') \infl P(\mu') \oplus P(\nu') \defl P(\omega).
 \]
 By Proposition~\ref{prop::exchange conflations are unique}, we get that~$P(\mu)\oplus P(\nu)$ and~$P(\mu')\oplus P(\nu')$ are isomorphic, and that the mutation conflations either both have the form~$P(\omega) \infl P(\mu) \oplus P(\nu) \defl P(\omega')$ or both have the form~$P(\omega') \infl P(\mu) \oplus P(\nu) \defl P(\omega)$.  Without loss of generality, assume that they have the first form.

 In particular, the leftmost morphisms of both mutation conflations are both minimal~$\add(P(\mu) \oplus P(\nu))$-approximations of~$P(\omega)$.  As such, they are isomorphic as morphisms.  They determine common substrings of~$\omega$ and~$\mu$ and of~$\omega$ and~$\nu$, which in turn determine~$\sigma$.  Point (ii) of Proposition~\ref{prop:exchangeablePairsNKC} is proved.
\end{proof}

@@

\subsubsection{A quotient of a Frobenius category}
\begin{example}
 Consider the preprojective algebra of type~$A_3$, as in Example~\ref{exam::preprojective silting objects}.  It is a self-injective algebra, so its module category is a Frobenius category.  Its projective-injective objects are the modules 
 \[
  \begin{matrix} 1 \\ 2 \\ 3 \end{matrix} \quad , \quad  \begin{matrix} 2 \\ 13 \\ 2 \end{matrix} \quad \textrm{and} \quad \begin{matrix} 3 \\ 2 \\ 1 \end{matrix}.
 \]
 Taking the quotient by the ideal generated by all morphisms factoring through direct sums of copies of the second of these objects, we get an extriangulated category whose Auslander--Reiten quiver is given by
  \[
  \hspace*{-2cm}
 {\footnotesize
  \begin{tikzcd}
   \begin{matrix} 3 \end{matrix}\arrow[dr] && \begin{matrix} 2 \\ 1 \end{matrix}\arrow[r]\arrow[dr] & \begin{matrix} 3 \\ 2 \\ 1 \end{matrix}\arrow[r]& \begin{matrix} 3 \\ 2 \end{matrix}\arrow[dr] && \begin{matrix} 1 \end{matrix}\arrow[dr]&& \begin{matrix} 2 \\ 3 \end{matrix} \\
    & \begin{matrix} 2\\ 13 \end{matrix}\arrow[ur]\arrow[dr] && \begin{matrix} 2 \end{matrix}\arrow[ur]\arrow[dr] && \begin{matrix} 13\\2 \end{matrix}\arrow[ur]\arrow[dr] && \begin{matrix} 2\\13 \end{matrix}\arrow[ur]\arrow[dr] &  \\
   \begin{matrix} 1 \end{matrix}\arrow[ur] && \begin{matrix} 2\\3 \end{matrix}\arrow[ur]\arrow[r] & \begin{matrix} 1\\2\\3 \end{matrix}\arrow[r] & \begin{matrix} 1\\2 \end{matrix}\arrow[ur] && \begin{matrix} 3\\1 \end{matrix}\arrow[ur] && \begin{matrix} 2\\1 \end{matrix}
  \end{tikzcd}
  }
 \]
 \pg{To finish}
\end{example}
